---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r load the data, echo=TRUE}
unzip("./activity.zip")
activity = read.csv("./activity.csv")
## remove rows that have steps = NA
cleaned<-na.omit(activity)
## convert date column to date type
cleaned$actualDate <- as.Date(as.character(cleaned$date), format = "%Y-%m-%d")
```

## What is mean total number of steps taken per day?

### Calculate the total number of steps taken per day
```{r q1, echo=TRUE}
aggregation <- aggregate(. ~date, data = cleaned, FUN = sum)
```
### Make a histogram of the total number of steps taken each day
```{r q1, echo=TRUE}
hist(aggregation$steps, main = "Histogram of steps taken per day", xlab="Steps taken per day")
```
### Calculate and report the mean and median of the total number of steps taken per day
```{r q1, echo=TRUE}
mean(aggregation$steps)
median(aggregation$steps)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
